CdGF_aucrf

### 2016-10-03 code/CdGF_AUCRF.Rmd
remove OTUs which select cage/inocula diffs

#random forest regression model to determine Day 0 OTU correlated with increased CFU
seed <- 1
n_trees <- 2001
iters <- 100

```

@@

Predict_df <- merge(meta_file[meta_file$day %in% 0, 
                              c('cage_id', 'mouse_id', 'Euth_Early')],
                    rel_abund_stool, by.x = 'mouse_id', by.y = 'row.names')

Predict_early_euth_df <- select(Predict_df, -cage_id, -mouse_id)

@@

LOcO_probs <- c()
for(i in 1:length(cage_list)){
  by_cage <- LOcO_df$cage_id == cage_list[i]
  LOcO_predict <- select(LOcO_df, -cage_id)
  train <- LOcO_predict[!by_cage,]
  test <- LOcO_predict[by_cage,]
  set.seed(j)
  temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
  LOcO_probs[LOcO_df$cage_id == cage_list[i]] <- predict(temp_model$RFopt, test, type='prob')[,2]
}
LOcO_roc <- roc(LOcO_df$Euth_Early~LOcO_probs)

#### plot rocs
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(otu_euth_roc, col='black', lwd=2, add=T, lty=1)
plot(cv10f_roc, col='gray', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=2)
plot(LOcO_roc, col='purple', lwd=2, add=T, lty=3)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('Leave-1 cage-out (AUC: %.3g)',LOcO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc),
                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(1,2,2,1,0,0), lwd=2, col=c('black','blue','purple','gray','red'), bty='n')
```

@@

Leave 1 cage out RF with all otus gives and AUC of 0.63-0.68, but using the reduced features selected by AUCRF on the whole dataset gives an AUC of 0.88-0.90

>>>>

#random forest regression model to determine Day 0 OTU correlated with increased CFU
seed <- 100316
n_trees <- 2001
iters <- 100
n_features <- 30
```

@@

Predict_df <- merge(meta_file[meta_file$day %in% 0, 
                              c('human_source', 'mouse_id', 'cage_id', 'Euth_Early')],
                    rel_abund_stool, by.x = 'mouse_id', by.y = 'row.names')
Predict_cage_df <- select(Predict_df, -human_source, -mouse_id, -Euth_Early)
Predict_source_df <- select(Predict_df, -cage_id, -mouse_id, -Euth_Early)
Predict_cage_df$cage_id <- factor(Predict_cage_df$cage_id)
Predict_source_df$human_source <- factor(Predict_source_df$human_source)

# identify OTUs that predict cage/inocula
set.seed(seed)
rf_cage <- randomForest(cage_id ~ ., data=Predict_cage_df, importance=TRUE, ntree=n_trees)
set.seed(seed)
rf_source <- randomForest(human_source ~ ., data=Predict_source_df, importance=TRUE, ntree=n_trees)
features_source <- head(sort(rf_source$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), n_features)
features_cage <- head(sort(rf_cage$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), n_features)
features_all <- unique(c(names(features_cage), names(features_source)))

# remove cage/inocula OTUs from prediction dataframes
Predict_early_euth_df <- select(Predict_df, -cage_id, -mouse_id, -human_source, -one_of(features_all))

@@

LOcO_probs <- c()
for(i in 1:length(cage_list)){
  by_cage <- LOcO_df$cage_id == cage_list[i]
  LOcO_predict <- select(LOcO_df, -cage_id)
  train <- LOcO_predict[!by_cage,]
  test <- LOcO_predict[by_cage,]
  set.seed(seed)
  temp_model <- AUCRF(Euth_Early~., data=train, pdel=0.99, ntree=500)
  LOcO_probs[LOcO_df$cage_id == cage_list[i]] <- predict(temp_model$RFopt, test, type='prob')[,2]
}
LOcO_roc <- roc(LOcO_df$Euth_Early~LOcO_probs)

#### plot rocs
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(otu_euth_roc, col='black', lwd=2, add=T, lty=1)
plot(cv10f_roc, col='gray', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=2)
plot(LOcO_roc, col='purple', lwd=2, add=T, lty=3)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('Leave-1 cage-out (AUC: %.3g)',LOcO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc) #,
#                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
#                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(1,2,2,1,0,0), lwd=2, col=c('black','blue','purple','gray','red'), bty='n')
```

@@

Leave 1 cage out RF with all otus gives and AUC of 0.70, but using the reduced features selected by AUCRF on the whole dataset gives an AUC of 0.80

To remove any cage effect RF could be picking out instead of the actual outcome, I first run RF on all OTUs to predict cage and separately human source. Then select the OTUS that are necessary for the lowest error rate (I went slightly above) from both cage and human souce RFs. Remove those OTUs from the OTU data frame and predict Euthanization with AUCRF
